---
title: "Dashboard"
format: 
 dashboard:
   html:
    fig-width: 10
    fig-height: 6 
    fig-align: center
   navbar: right
   css: dashboard.scss
execute:
  echo: false
  warning: false
---

```{r local_packages}
#| include: false
```

{{< include ../_setup.qmd >}}

# World Population {orientation="columns"}

##  Column

### Row {height=20%}
```{r}
#| content: valuebox
#| title: "2000 World Population"

list(
  icon = "globe-europe-africa",
  color = "secondary",
  value = paste(pop_world(year=2000), " B")
)
```

```{r}
#| content: valuebox
#| title: "2024 World Population"

list(
  icon = "globe-central-south-asia",
  color = "secondary",
  value = paste(pop_world(), " B")
)
```

### Row {height=28%}
::: {.card}
The United Nations forecasts that the worldwide population will peak before the end of the century, surpassing `r pop_world(year=2100)` billion individuals. Contrary to widespread belief, the world's population is not undergoing exponential growth. Although the total number of people worldwide continues to rise, the rate of population growth reached its maximum decades ago. To maintain a stable population, the Total Fertility Rate (TFR) must be at replacement level (2.1 children per woman). The global convergence of total fertility rates (TFR) at or below the replacement level will eventually reduce overall population sizes to earlier levels. However, the demographic composition at that point will differ significantly from the past. The world will face higher old-age dependency ratios, raising concerns and likely **intensifying competition for young migrants**.
:::

### Row {height=50%}
```{r}
# 1. Long Run Population Growth (OWID)
world_long_run_owid
```

## Column

```{r}
# 2. Population Size Evolution
world_pop_size
```

```{r}
# 3. Fertility Rates
fertility_rates_decline
```

# Fertility Rates {orientation="columns"}

## Column 

### Row {.tabset}

```{r}
#| title: "Africa Doubling"
# 4. Africa Doubling Population
africa_double_2060
```

```{r}
#| title: "Africa TFR Convergence"
# 5. Declining Fertility Convergence
africa_tfr_decline
```

### Row

```{r}
# 14. TFR distribution boxplot
tfr_regions_boxplots
```

## Column

```{r}
#| fig-width: 9
#| fig-height: 9
# 7. Regional Dependency Ratios
region_dep_ratio &
  plot_annotation(theme = theme(plot.background = element_rect(colour="white")))
```

# Growth Momentum

## Row

## How population momentum works

```{r}
# 11. How Population Momentum Works
dem_momentum1
```

```{r}
# 11. How Population Momentum Works
dem_momentum2
```

```{r}
# 11. How Population Momentum Works
dem_momentum3
```
## Row 

```{r}
# 12. TFR Cumulative Plots: Regions
cum_tfr_regions
```

```{r}
# 13. TFR Cumulative Plots: Sub-Regions
cum_tfr_subregions
```

# Working-age Population {orientation="columns"}

## Column

### Row

```{r}
# 6. Working-age Population Annual Change
wap_annual_change
```

### Row {.tabset} 

```{r}
#| title: Size Chanpions
#| fig-width: 10
# 15. Working-age population size chanpions VS TFR champions

# Working-age population size champions
twa_size_champions
```

```{r}
#| title: TFR Chanpions
#| fig-width: 10
# TFR champions
twa_tfr_champions
```

## Column 

### Row {height=10%}

Between 2024 and 2060, Africa's working-age population is projected to increase by about <b>`r dc |> pull(change) |> round()` percent</b>, equating to roughly <b>`r dc |> pull(diff) |> round()` million people</b>.

### Row {.tabset}

```{r}
#| title: Regions 
#| fig-height: 9
#| fig-width: 10
# 17. Working-age Distribution ridgeplot 
# Regions
wap_ridge_regions + labs(
  title = "Regions Working-age Population Distribution"
)
```

```{r}
#| title: Sub-regions
#| fig-height: 9
#| fig-width: 10
#Subregions
wap_ridge_subregions + labs(
  title = "Sub-regions Working-age Population Distribution"
)
```

# Dependency Ratios {orientation="columns"}

## Column

```{r}
# 10. Africa being left behing
left_behind
```

```{r echo=FALSE, warning=FALSE, out.width="100%"}
# 8. Old-age VS child dependency ratios
old_child_dep_ratios
```

## Column

```{r}
#| fig-width: 12
#| fig-height: 13
# 9. School-age population evolution
school_age_pop
```



